---
title: 'Master R Developer Workshop: Functional Programming'
author: "Mohit Bhasin"
date: "May 27, 2015"
output: html_document
---

#DRY
Functions first look inside their lexical scope, then looks at global 
Use return(), only if you want to return something early, otherwise R automatically returns the last line
This is true of most functional programming languages
```{r}

x <- 5 
f <- function() {  
  x <- 20
  y <- 10
  c(x = x, y = y) }
f() 

x <- 5 
f <- function() {  
  y <- 10
  c(x = x, y = y) }
f() 

```

Understanding functional environments: They vanish after excution

```{r}

j <- function()
  {   if (!exists("a")) {
    a <- 5   } 
  else {
    a <- a + 1   }
  print(a) }
j()
j()

```

#Functionals
One of the arguments of the function is another function 
```{r}
set.seed(1014) 
# Create some random output: # 20 random vectors with random lengths 
l <- replicate(3, runif(sample(1:10, 1)), simplify = FALSE)
str(l) 
l

```


Use ```seq_along()``` instead of ```1:length(x)```
```{r}
# Extract length of each element 

#always specify the length of an array in a for loop in advance, extending it as you progress through the loop is inefficient
lengths <- vector("list", length(l))

for (i in seq_along(l)) {
  lengths[[i]] <- length(l[[i]]) 
  }
lengths

```


#Design your first functional
Abstract over forloops
start by taking a specific solution that works and then generalize it
```{r}
# ... is a placeholder
do_each <- function (x,f,...){
  
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]],...) 
  }
  out
}

do_each(l,mean)

```
We just built lapply

The function below scales a vector so it falls in the range [0, 1]. How would you apply it to every column of a data frame?

```{r}
scale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)   
  (x - rng[1]) / (rng[2] - rng[1]) 
}
```

Solution 1
```{r}
for (i in seq_along(mtcars)){
  mtcars[[i]] <- scale01(mtcars[[i]])
}
#rm(mtcars)

```

Solution 2
```{r}
mtcars <- lapply(mtcars, scale01) 
#mtcars # a list :( 
rm(mtcars) 
mtcars[] <- lapply(mtcars, scale01)  # NEAT
#mtcars # a data frame :)
rm(mtcars)
```

vapply always gives you a logical vector, sapply output depends on input. sapply is a wrapper around lapply. Never use sapply inside a function  
<bit.ly/advr-chicago>

#Function Operators
They abstract away common uses of anonymous functions
We will be skipping this, not used often
```{r}
library(pryr) 
sapply(mtcars, function(x) length(unique(x)))
sapply(mtcars, compose(length, unique))

```

```{r}
x <- 0
y <- 10
f <- function() {   
  x <- 1
  function() { 
    y <- 2
    x + y
  }
}
f()
f()()
```

#Function Factories
Functions can generate functions. A list can contain functions.
<http://adv-r.had.co.nz/Functions.html>
<http://adv-r.had.co.nz/Functional-programming.html>
<http://adv-r.had.co.nz/Functionals.html>
<http://adv-r.had.co.nz/Function-operators.html>
```{r}
# Closures are useful when you want a function 
# that can create a whole class of functions:
# a function factory
power <- function(exponent) {
  function(x) {
    x ^ exponent  }}
square <- power(2)
square(2)
square(4)
cube <- power(3)
cube(2)
cube(4)

square # We can find the environment and its parent
environment(square)
as.list(environment(square))
as.list(environment(cube))

pryr::unenclose(square)
pryr::unenclose(cube)

```


#APPENDIX
List of cool functions
```{r}
?simplify2array() #Base R function that tries to coerce a list into an array or atrix
as.list(environment("f"))
```



