---
title: "Master R Developer Workshop: Object Oriented Programming"
author: "Mohit Bhasin"
date: "May 27, 2015"
output: html_document
---

How do you see the source code of the %in% function? What about diag<-?If you have a column named "Income ($1000)" in your data frame called df, how can you access it with $?

When using uncommon variable names use backquotes
```{r}
`%in%`
`diag<-`

```

#Attributes
```{r}
x <- 1:6
attr(x, "max") <- 5
attr(x, "max")
attributes(x)

```
What are factors made of? Ans: Integers
```{r}
f <- factor(c("a","b"))
typeof(f)
attributes(f)

```
What are dataframes made of? Ans: Lists
```{r}
df <- data.frame(x=2, y=3)
typeof(df)
attributes(df)

```

#Classes
Say we wanted to stroe a discrete random variable, how would you store them and what would they need?
```{r}
x <- c(-1,0,1)
p <- c(0.3, 0.3, 0.4)

# Constraints
all(p > 0) 
sum(p) == 1 
abs(sum(p) - 1) < 1e-6 

# Ways to store 
structure(x, prob = p) 
structure(p, val = x) 
list(x = x, p = p)

# Start by defining a constructor function. It
# uses structure to set the class attribute.
rv <- function(x, probs = NULL) {   
  if (is.null(probs)) {
    probs <- rep(1, length(x)) / length(x)   
    }
  structure(x, probs = probs, class = "rv") } 

# Also customary to create function to test if # an object is of that class: 
is.rv <- function(x) {    inherits(x, "rv")  
  # equivalent to "rv" %in% class(x) 
}

# And we'll also write a helper to extract probabilities
probs <- function(x) attr(x, "probs")

```

What's wrong with the following objects? 
```{r}
rv(1:3, c(-1, 2)) 
rv(c(1, 1), c(0.5, 0.5))
```

In R methods belong to functions not to classes or objects so you dont see rv.mean()
```{r}
tail(methods(class = "factor"))
tail(methods(class ="data.frame"),5)

```

recommends not using "." in naming variables, except when it is actually associated with the object
Usually start by building a print method for 
```{r}
print.rv <- function(x,..){
  cat("values = ", x ,"\n", "probs =" , probs(a))
}

a = rv(1:4)
a # now same as print_rv(a)
```

Create a mean method
```{r}
mean.rv <- function (x,...){
  cat("E[x] =", sum(probs(x)*x))
}
mean(a) # we have now overwritten the default way mean works and applied it to behve correctly for RV. 
```

#Inheritance
Creating the subsetting method for rvs
```{r}
`[.rv` <- function(x, i, ...) {
  rv(NextMethod(), prop.table(probs(x)[i])) # NextMethod ends up calling the default method
}
a[1:2]

```


