---
title: "Master R Developer Workshop: Writing Good Functions "
author: "Mohit Bhasin"
date: "May 27, 2015"
output: html_document
---

Hadley Wickham
Chief Scientist
Rstudio
Wednesday, May 27, 2015 at 9:00 AM - Thursday, May 28, 2015 at 5:00 PM
Chicago, IL

#Debugging
Use the interactive debugger, it calls setbreakpoint() which is essentially the same as calling browser()
"Q" to quit
```{r}
browser()

```

#Robust code
Be explicit, use TRUE, FALSE, not T,F
Avoid functions like ```(subset, with, transform)``` because of poor error messages and eccentric behavior. likewise for ```sapply```

#Practice problem
Fix this function

```{r}
col_means <- function(df) {   
  numeric <- sapply(df, is.numeric)   
  numeric_cols <- df[, numeric]   
  data.frame(lapply(numeric_cols, mean)) 
  }

head(mtcars)
col_means(mtcars)
# col_means(mtcars[,0]) Breaks knitr
# col_means(mtcars[0,])
# col_means(1:10)

```

sapply - sometimes give you a logical vector, sometimes it gives you a list  
What should the output be? Ans: always be a df with one row, the colmeans of each col  
There is no rule about strict Versus helpful, i.e how should we handle matrix vs data.frame  
The worst thing a function can do is give the wrong output silently  

```{r}

col_means <- function(df) {   
  ## CALCULATES THE MEANS OF EACH COLUMN OF A DATA FRAME
  stopifnot(is.data.frame(df))     
  numeric <- vapply(df, is.numeric, logical(1))
  # You always want to do drop = FALSE
  numeric_cols <- df[, numeric, drop = FALSE]   
  data.frame(lapply(numeric_cols, mean)) 
  }

```

Imagine you have a vector of events that you # want to divide into groups. You know when # an event ends. How could you generate a  # unique integer for each group? 
```{r}
x <- sample(c(TRUE, FALSE), prob = c(0.2, 0.8), 100, rep = T) # Brainstorm for 2 minutes
```

Solution 1: Too clever?
```{r}
cumsum(x) + 1 #incomplete solution needs some test
```

Solution ?
```{r}
which (x)
```

#Commenting
Use comments to answer why, not how or what.  
<http://adv-r.had.co.nz/Style.html>

#Profiling Code
```{r}
library(microbenchmark)
#install.packages("multcomp")
x <- runif(100) 
microbenchmark(
  sqrt(x),
  x ^ 0.5,   
  exp(log(x) / 2) 
  )
```

Your turn  
What's the fastest way to get the bottom-right value from mtcars (i.e. 32nd from top, 11th from left)? e.g. mtcars[32, 11]

```{r}
library(microbenchmark)
microbenchmark(   
  mtcars["Volvo 142E", "carb"],
  mtcars[32, 11],
  # A DATA FRAME IS A LIST OF VECTORS
  mtcars[[c(11, 32)]],
  mtcars[["carb"]][32],
  mtcars[[11]][32], # EQUIVALENT TO mtcars[[11]][[32]]
  mtcars$carb[32],
  .subset2(mtcars, 11)[32] 
)
```
